---
title: "COMPSCIX 415.2 Homework 5/Midterm"
author: "Lionel Armstrong"
date: "February 28, 2018"
output:
  html_document:
    toc: true
    theme: united
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr) 
library(tidyverse)
library(dbplyr)
library(ggplot2)
```
https://github.com/lionelja/compscix-415-2-assignments


#RStudio and R Markdown (3 points)  


*1.Use markdown headers in your document to clearly separate each midterm question and add a table of contents to your document.*  


#The tidyverse packages (3 points)  


By now you’ve used at least five different packages from the tidyverse for plotting, data munging, reshaping data, importing/exporting data, and using tibbles (the tibble package is used for this without you even realizing it’s there).

*1.  Can you name which package is associated with each task below?*

 Plotting -
 Data munging/wrangling -
 Reshaping (speading and gathering) data -
 Importing/exporting data -*
 
#####ANSWER:#####
The ggplot2 package is used for plotting data.
The dplyr package is used for data mungung/wrangling.
The tidyr package is used for reshaping data.
The readr package is used for importing/exporting data.

 
*2.  Now can you name two functions that you’ve used from each package that you listed above for these tasks?*  

Plotting -
 Data munging/wrangling -
 Reshaping data -
 Importing/exporting data (note that readRDS and saveRDS are base R functions) -

#####ANSWER:#####
Two ggplot2 functions I've used are geom_smooth() and geom_bar().
Two dplyr functions I've used are select() and filter().
Two tidyr funtions I've used are gather() and spread().
Two readr functions I've used are read_csv() and write_csv().

#R Basics (1.5 points)    

*1.  Fix this code with the fewest number of changes possible so it works:*  
My_data.name___is.too00ooLong! <- c( 1 , 2   , 3 )

#####ANSWER:#####
Underscores and periods are the only valid special characters in a variable. Removing the explanation point make the statement valid.

```{r}
My_data.name___is.too00ooLong <- c( 1 , 2   , 3 )
```

*2.  Fix this code so it works:*  
my_string <- C('has', 'an', 'error', 'in', 'it')

#####ANSWER:#####
The components of a vector is designated with a capital "C" isn't valid. This should be a lower case "c".

```{r}
my_string <- c('has','an','error','in', 'it')

```

*3.  Look at the code below and comment on what happened to the values in the vector.* 

my_vector <- c(1, 2, '3', '4', 5)
my_vector
 ## [1] "1" "2" "3" "4" "5"
 
#####ANSWER:#####
The components of a vector must be all of the same type. When the character type component '3' is encounter all components become charater type.
 
#Data import/export (3 points)  


*1.  Download the rail_trail.txt file from Canvas (in the Midterm Exam section here) and successfully import it into R. Prove that it was imported successfully by including your import code and taking a glimpse of the result.*  

#####ANSWER:#####
```{r}
myfile='C:/Rdata/rail_trail.txt'
my_data <- read_delim(file = myfile, delim='|')
glimpse(my_data)
```

*2.  Export the file into an R-specific format and name it “rail_trail.rds”. Make sure you define the path correctly so that you know where it gets saved. Then reload the file. Include your export and import code and take another glimpse.*  

#####ANSWER:#####
```{r}
saveRDS(my_data, file='c:/RData/rail_trail.rds')
rds_file='C:/Rdata/rail_trail.rds'
my_rds_data <- readRDS ('C:/Rdata/rail_trail.rds')
glimpse(my_rds_data)
```

#Visualization (6 points)  
*1.  Critique this graphic: give only three examples of what is wrong with this graphic. Be concise.*  

#####ANSWER:#####
The bubble size is at the lower end of the spectum of good visual cues. The age deminsion is omitted for men and women graph.  The numeric values in the center of the buble isn't clear whether not it's an percentage for two reason (1) not labeled with %, (2) the total percentage value for a nomial deminsion such as sex should equal 100%.


*2.  Reproduce this graphic using the diamonds data set.*  

#####ANSWER:#####

```{r}
ggplot(data = diamonds, mapping = aes(y = carat, x = cut, fill= color)) +
  xlab('CART OF DIAMONDS') + ylab('CUT OF DIAMONDS') + 
  geom_boxplot(position = 'identity') +
  coord_flip()
```


*3.  The previous graphic is not very useful. We can make it much more useful by changing one thing about it. Make the change and plot it again.*  

#####ANSWER:#####

```{r}
ggplot(data = diamonds, mapping = aes(y = carat, x = cut, fill= color)) +
  xlab('CART OF DIAMONDS') + ylab('CUT OF DIAMONDS') + 
  geom_boxplot(position = 'identity') +
  coord_flip() +
   facet_wrap(~ color)
```

#Data munging and wrangling (6 points)  

*1.  Is this data “tidy”? If yes, leave it alone and go to the next problem. If no, make it tidy. Note: this data set is called table2 and is available in the tidyverse package. It should be ready for you to use after you’ve loaded the tidyverse package.*  

#####ANSWER:#####

table2 is tidy.  Each row is a type of count for given year and each column has the same type of value in each row. 

```{r}
table2
```

*2.  Create a new column in the diamonds data set called price_per_carat that shows the price of each diamond per carat (hint: divide). Only show me the code, not the output.*  

#####ANSWER:#####

```{r}
my_diamonds <- diamonds %>% mutate(price_per_carat = carat / price )
```

*3.  For each cut of diamond in the diamonds data set, how many diamonds, and what proportion, have a price > 10000 and a carat < 1.5? There are several ways to get to an answer, but your solution must use the data wrangling verbs from the tidyverse in order to get credit.*

    ◦Do the results make sense? Why?
    ◦Do we need to be wary of any of these numbers? Why?

#####ANSWER:#####

```{r}
my_diamonds <- diamonds %>% mutate(price_per_carat = carat / price )
```

#EDA (6 points)  

Take a look at the txhousing data set that is included with the ggplot2 package and answer these questions:

*1.  During what time period is this data from?*  
#####ANSWER:#####

The txhousing data is from the time period beginning January 2001 and ending July 2015. 

```{r}
end_period <- txhousing %>% arrange(desc(year), desc(month))
end_period
begin_period <- txhousing %>% arrange(year, month)
begin_period
```

*2.  How many cities are represented?* 

#####ANSWER:#####

There are forty-six cities represented in the txhousing data. 

```{r}
Q2_answer <- txhousing %>% summarize(num_cities = n_distinct(city, na.rm=TRUE))
Q2_answer
```

*3.  Which city, month and year had the highest number of sales?*  

#####ANSWER:#####

The highest number of sales in txhousing data was recorded in July 2015 in the city of Houston. data. 

```{r}
maxsale <- txhousing %>% arrange(desc(sales))
maxsale 
```

*4.  What kind of relationship do you think exists between the number of listings and the number of sales? Check your assumption and show your work.* 

*5.  What proportion of sales is missing for each city?*  

*6.  Looking at only the cities and months with greater than 500 sales: ◦Are the distributions of the median sales price (column name median), when grouped by city, different? The same? Show your work.*
   ◦Any cities that stand out that you’d want to investigate further?
   ◦Why might we want to filter out all cities and months with sales less than 500?

#Git and Github (1.5 points)  

*To demonstrate your use of git and Github, at the top of your document put a hyperlink to your Github repository.*

Once you are finished with your midterm, commit your final changes with the comment “finished the midterm - woohoo” and push your R Markdown file and your html or pdf file to Github.






